---
title: "Course Project 1: Simulation Exercise"
author: "Xiaozhu Zhang"
date: "Jan. 13, 2021"
output: pdf_document
---

In this part I will investigate the exponential distribution and compare it with CLT. If a random variable $X\sim \text{Exp}(\lambda)$, then
\[ f(x) = \lambda e^{-\lambda}, \quad \mathbb{E}(X) = \frac{1}{\lambda}, \quad {\rm Var}(X) = \frac{1}{\lambda^2}.   \]

Now I will set $\lambda=0.2$, and investigate the distribtuion of averages of 40 exponentials (namely $\bar{X} = (\sum_{i=1}^{40}X_i)/40$, for i.i.d. $X_i$) with 1000 simulations. 

## 1. Simulation process
First, set the parameter $\lambda$, the sample size $n$, and the number of simulations $N$.
```{r}
lambda = 0.2
n = 40
N = 1000
```

Then sample $40*1000$ values from the $\text{Exp}(0.2)$ density, store them in a matrix, and use the ```apply``` function to obtain the ```Sample``` of length 1000.
```{r}
set.seed(1234)
Sample <- matrix(rexp(n = n * N, lambda), N, n)
Sample <- apply(Sample, 1, mean)
```

## 2. Mean comparison
We can draw the histogram of ```Sample```, where the blue line, $\mathbb{E}(\bar{X}) = 1/\lambda$, is the theoretical mean, while the red line is the sample mean of the distribution. It is clear from figure 1 that they are very close to each other. The exact numbers shown below give the same result.
```{r}
# Theoretical mean
1/lambda
# Sample mean
mean(Sample)
```
 
```{r, warning=FALSE, echo=FALSE, fig.height=2, fig.width=4, fig.align='center', fig.cap="Histogram of 1000 averages over 40 exponentials"}
library(ggplot2)
ggplot(data.frame(Sample), aes(x = Sample)) + 
    geom_histogram(colour = "black", fill = "white", bins = 30) +
    geom_vline(aes(xintercept = 1/lambda),
               color = "blue", linetype = "dashed", size = 0.9) +
    geom_vline(aes(xintercept = mean(Sample)),
               color = "red", linetype = "dashed", size = 0.9) + 
    labs(title = "Histogram of 1000 averages", x = "x")

```

## 3. Variance Comparison
The theoretical variance is ${\rm Var}(\bar{X}) = {\rm Var}(X)/n = 1/(n\lambda^2)$. 
```{r}
# Theoretical variance
1/(n * lambda^2)
# Sample variance
var(Sample)
```
Still, they are close. But in a way the difference between theoretical and sample variance is beyond nuance, implying that $n=40$ may not be large enough since CLT requires a large sample size. For example, if we increase $n$ to 100 or 150, then the gap between theoretical and sample variance will decrease, and from the figure below we conclude that the distribution will be more centered.
```{r, echo = FALSE, fig.cap="Histogram on different sample sizes", fig.height=2.4, fig.width=4.5, fig.align='center'}
SD <- function(lambda, n, N){
  set.seed(1234)
  Sample <- matrix(rexp(n = n*N, lambda), N, n)
  Sample <- apply(Sample, 1, mean)
  return(list(Sample = Sample))
}
Sp0 <- Sample
Sp1 <- SD(lambda, 100, N)$Sample
Sp2 <- SD(lambda, 150, N)$Sample
Sp <- c(Sp0, Sp1, Sp2)
label <- factor(rep(c("40","100","150"), each = 1000), 
                levels = c("40","100","150"), ordered = TRUE)

theory_var <- 1/(c(40,100,150)*lambda^2)
sample_var <- c(var(Sp0),var(Sp1),var(Sp2))
dat <- data.frame(theory_var, sample_var)
rownames(dat) <- c("n=40","n=100","n=150")
colnames(dat) <- c("theory_variance", "sample_variance")
dat

ggplot(data.frame(Sp, size = label), aes(x = Sp)) + 
  geom_histogram(aes(fill = size), bins = 30) +
  scale_fill_manual(values=alpha(c("#6495ED","#FFA500","#FF4500"), 0.7)) +
  facet_grid(.~ size) +
  labs(title = "Histogram on different sample sizes", x="x",y="count")
```


## 4. Distribution
The sampling distribution of $\bar{X}_n$ is shown in figure 1, while the population distribution is shown in figure 3. The distribution from population to sampling is becoming more symmetric and more like normal.
```{r, echo=FALSE, fig.cap="Histogram of population", fig.height=2, fig.width=4, fig.align='center'}
dat <- data.frame(x = rexp(n = 1000, lambda))
ggplot(dat, aes(x = x)) + 
  geom_histogram(colour = "black", fill = "white", bins = 30) +
  labs(title = "Histogram of population", x="x",y="count")
```
In order to investigate more about the normality, we can draw the qq-plot as figure 4. We conclude that the sampling distribution is *approximately normal* although with few outlier points indicating a little skewness. 
```{r, message=FALSE, fig.cap="qq-plot of sampling distribution", fig.align='center', fig.height=2, fig.width=4}
library("ggpubr")
ggqqplot(Sample)
```

In a nutshell, as $n$ goes to infinity, we have
\[  \bar{X} \rightarrow N(\frac{1}{\lambda}, \frac{1}{n\lambda^2})  \] in distribution.

\newpage
# Appendix
```{r, eval=FALSE}
# Codes for figure 1
library(ggplot2)
ggplot(data.frame(Sample), aes(x = Sample)) + 
  geom_histogram(colour = "black", fill = "white", bins = 30) +
  geom_vline(aes(xintercept = 1/lambda),
             color = "blue", linetype = "dashed", size = 0.9) +
  geom_vline(aes(xintercept = mean(Sample)),
             color = "red", linetype = "dashed", size = 0.9) + 
  labs(title = "Histogram of 1000 averages", x = "x")
```
  
```{r, eval=FALSE}
# Codes for figure 2 and the table
SD <- function(lambda, n, N){
  set.seed(1234)
  Sample <- matrix(rexp(n = n * N, lambda), N, n)
  Sample <- apply(Sample, 1, mean)
  return(list(Sample = Sample))
}
Sp0 <- Sample
Sp1 <- SD(lambda, 100, N)$Sample
Sp2 <- SD(lambda, 150, N)$Sample
Sp <- c(Sp0, Sp1, Sp2)
label <- factor(rep(c("40", "100", "150"), each = 1000), 
                levels = c("40" ,"100", "150"), ordered = TRUE)
# the table
theory_var <- 1/(c(40, 100, 150) * lambda^2)
sample_var <- c(var(Sp0), var(Sp1), var(Sp2))
dat <- data.frame(theory_var, sample_var)
rownames(dat) <- c("n=40", "n=100", "n=150")
colnames(dat) <- c("theory_variance", "sample_variance")
dat

# figure 2
ggplot(data.frame(Sp, size = label), aes(x = Sp)) + 
  geom_histogram(aes(fill = size), bins = 30) +
  scale_fill_manual(values = alpha(c("#6495ED", "#FFA500", "#FF4500"), 0.7)) +
  facet_grid(.~ size) +
  labs(title = "Histogram on different sample sizes", x = "x", y = "count")
```
  
```{r, eval=FALSE}
# Codes for figure 3
dat <- data.frame(x = rexp(n = 1000, lambda))
ggplot(dat, aes(x = x)) + 
  geom_histogram(colour = "black", fill = "white", bins = 30) +
  labs(title = "Histogram of population", x = "x", y = "count")
```


