---
title: "Demo of slides"
author: "Paige Zhang"
date: "Feb. 25, 2021"
output: 
  beamer_presentation:
    theme: "Madrid"
    colortheme: "default"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

-   Bullet 1
-   Bullet 2
-   Bullet 3

And there are others:

1.  California

2.  New York

3.  North Carolina

## Slide with R Output

```{=tex}
\begin{theorem}[Newton-Einstein]
This is ...
\end{theorem}
```
```{r cars, echo = TRUE}
s <- summary(cars)
knitr::kable(s, aligh = 'c', caption = "The Summary")
```

## Slide with Plot

```{r pressure}
library(ggplot2)
ggplot(data = mtcars, aes(x = mpg)) + geom_histogram(bins = 30) 
```
