---
title: "Solution to Quiz 1"
output: html_notebook
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Question 1
Q: The function `ggplot()` requires that the value supplied to the 'data' argument be a

A: data frame

## Question 2
Q: You run the following code and get the error printed directly after the code: 
```{r}
library(ggplot2) 
library(faraway) 
data(nepali) 
head(nepali, 3) 

## id sex wt ht mage lit died alive age 
## 1 120011 1 12.8 91.2 35 0 2 5 41 
## 2 120011 1 12.8 93.9 35 0 2 5 45 
## 3 120011 1 13.1 95.2 35 0 2 5 

ggplot(nepali, aes(x = ht, y = wt)) 
+ geom_point() 

## Error in +geom_point(): invalid argument to unary operator 
```
What happened?

A: Because you put the + at the start of the second line, instead of at the end of the first line, R thought
the call was over and tried to run the first line by itself, and then tried to run the second line by itself,
without an associated ggplot object.


## Question 3
Q: You have a dataset of observations on study subjects which has height and weight for each subject measured at multiple time points. You want to create a scatterplot of height (x-axis) by weight (y-axis). Since there are multiple measurements per subject, you would like to plot the measurements for each subject in a separate color. There is a column in the data called id that gives the unique id of each study subject; this column currently has the class "numeric". Which of the following choices is a good strategy for creating a plot where data points use color to identify the subject? 

A: Convert the `id` column of the data frame to a factor, so a discrete color palette will be used, then use `ggplot2` to create a scatterplot and specify `color = id` in the `aes()` section of the `ggplot()` call.

Example code:
```{r}
df %>% 
  mutate(id = factor(id)) %>%
  ggplot(aes(x = height, y = weight, color = id)) + 
  geom_point()
```


## Question 4
Q: You want to save a figure from R as a pdf file named "MyFig.pdf". Which of the following describes the proper steps to take?

A: First run `pdf("MyFig.pdf")` to open a pdf device, then run all your code to create the figure, then use `dev.off()` to close the device.

Example code:
```{r}
pdf("MyFig.pdf")
ggplot(df, aes(x = x, y = y)) + 
  geom_point()
dev.off()
```


## Question 5
Q: Consider the following dataset.
```{r}
library(lattice)
data(barley)
head(barley)

##     yield   variety year            site
## 1 27.00000 Manchuria 1931 University Farm
## 2 48.86667 Manchuria 1931          Waseca
## 3 27.43334 Manchuria 1931          Morris
```

Which ggplot2 expression would make a scatterplot of year and yield while coloring each point according to its variety?

A: `ggplot(data = barley, aes(year, yield)) + geom_point(aes(color = variety))`


## Question 6
Q: Consider the following dataset.
```{r}
data(trees)
head(trees)

##  Girth Height Volume
##   8.3     70   10.3
##   8.6     65   10.3
##   8.8     63   10.2
##  10.5     72   16.4
##  10.7     81   18.8
##  10.8     83   19.7
```
You want to create a scatter plot showing `Girth` on the x-axis, `Height` on the y-axis, and `Volume` with point size. Which of the following code would create that plot?

A: (1)
```{r}
trees %>%
  ggplot(aes(x = Girth, y = Height)) +
  geom_point(aes(size = Volume))
```

(2)
```{r}
trees %>%
  ggplot(aes(x = Girth, y = Height, size = Volume)) +
  geom_point()
```

(3)
```{r}
ggplot() +
  geom_point(data = trees,
             aes(x = Girth, y = Height, size = Volume))
```

(6)
```{r}
trees %>%
  ggplot() +
  geom_point(aes(x = Girth, y = Height, size = Volume))
```

(8)
```{r}
ggplot(trees, aes(x = Girth, y = Height, size = Volume)) +
  geom_point()
```


## Question 7
Q: How can you create small multiples using ggplot2?

A: (1) Add `facet_wrap` to the geom, specifying the formula to use for faceting

(2) Add `facet_grid` to the geom, specifying one or two columns to use for faceting


## Question 8
Q: What does `geom_smooth(method = "lm")` do when it is added to a ggplot object?

A: It adds a fitted linear regression line to the plot


## Question 9
Q: When plotting multiple panels using faceting for a single variable, what controls the order in which the panels are plotted?

A: Panels are ordered by the alphabetical order or by the order of the levels of the faceting variable

## Question 10
Q: What function would you use to limit the range of the x-axis in a scatterplot?

A: `scale_x_continuous()`